Storm surge modelling in the Strait of Georgia

### 2014-07-26 RoughWork/manuscript.tex
Included description of nove 2009

\subsection{Strong Wind Event: Nov 18, 2009}

@@

\subsection{Statisitics}
%I think Hal Ritchie's talk with the storm surges had some stats presented. Look this up...

>>>>

\subsection{Strong Wind Event: Nov 19, 2009}
Next, a relatively strong wind event recorded at Point Atkinson on Nov 19, 2009 is considered. As recorded in the Environment Canada Historical Climate Data archives, the maximum hourly wind speed at Popint Atkinson was 18.6 m/s and the largest gust observed was 25.8 m/s or 93 km/hr. In addtion, large wind speeds over 15 m/s persisted for over 9 hours. Yet, the surge associated with this event is rather small when compared the previous two cases as shown in Figure \ref{fig:nov2009} where we have plotted the wind speed at Point Atkinson, sea surface height anomaly forcing at the open boundaries, and observed and modelled residuals at Point Atkinson. First, the largest surge throughoutt this time frame did not coincide with the large wind event, reflected in both the model and observations. The observed surge reached 62.6 cm on Nov 20 at 10:00 hr, while the model recorded a maximum surge of 38.8 cm five hours later. The surge did not coincide with a high tide and so no high water level events were recorded. Although not shown here, the other three locations also observed a maximum surge on Nov 20, which appears to be linked to a secondary high wind event and a more siginificant surge entering the system at Tofino and Port Hardy.

\citet{danard2003storm} discuss the role of wind fetch, the distance over which wind acts on the surface of a water basin, in setting up seiche motion in lakes. As wind blows over the surface of a lake, the water level becomes elevated on the downwind side and depressed on the upwind side. The longer the fetch, the more significant the tilting of the water surface. These changes in water level can also contribute to storm surges.  During the high wind event on Nov 19, the winds at Point Atkinson were directed to the west, explaining the lack of surge associated with this high wind event since wind stresses would act to depress the water on the east side of the basin. Perhaps the small surge that occured after the high wind event has formed in response to the sea surface returning to equilibrium in a seiche-like motion after the winds have relaxed. The model appears to reproduce this surge with reasonable accuracy, however, there are some discrepencies between weather observations and the atmospheric forcing in use. A more detailed look at the senstivity of modelled surges to atmospheric forcing is considered in a case using a high-resolution weather product.

\begin{figure}
\centering
\includegraphics[scale=0.6]{Figures/nov2009_bg.pdf}
\caption{Wind and storm surge conditions at Point Atkison from  strong wind event on Nov 19, 2009. Top: Wind speed from Environment Canada observations at Point Atkinson (solid) and and from the atmophseric model (CGRF) at a grid point near Point Atkinson (dashed). Middle: Sea surace height anomaly at Tofino (solid) and Port Hardy (dashed) used as forcing conditions at the western and northern open boundaries. Bottom: Observed (solid) and modelled (dashed) residuals at Point Atkinson.  }
\label{fig:nov2009}
\end{figure}

@@

\subsection{Statistical evaluation}
%I think Hal Ritchie's talk with the storm surges had some stats presented. Look this up...
